---
title: "Home Cage Sipper — Circadian Drinking Analysis"
author: "Rachel C. Rice"
date: "June 23rd, 2026"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Overview

This document analyzes circadian drinking rhythms from a two-bottle choice
home-cage sipper experiment.

Factors:

- Fluid group: EtOH (ethanol + water) or H2O (water only)
- Sex
- Age: 3 weeks, 6 weeks, 18 weeks

---

# Setup

## Libraries

```{r libraries}
library(ggplot2)
library(patchwork)
library(plyr)
library(dplyr)
library(ez)
library(stats)
```

## summarySE helper function

Source: http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/
  Custom function to call mean +/- SEM

```{r summarySE}
summarySE <- function(data = NULL, measurevar, groupvars = NULL,
                      na.rm = FALSE, conf.interval = .95, .drop = TRUE) {
  length2 <- function(x, na.rm = FALSE) {
    if (na.rm) sum(!is.na(x)) else length(x)
  }
  datac <- ddply(data, groupvars, .drop = .drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm = na.rm),
                     mean = mean(xx[[col]], na.rm = na.rm),
                     sd   = sd(xx[[col]], na.rm = na.rm))
                 }, measurevar)
  datac$se <- datac$sd / sqrt(datac$N)
  ciMult   <- qt(conf.interval / 2 + .5, datac$N - 1)
  datac$ci <- datac$se * ciMult
  return(datac)
}
```

---

## Time-lookup table

Build a sequential time index starting at lights off
  10:00 (lights off) = Time_Number 0
  22:00 (lights on) = Time_Number 720

Any rows recorded before 10 AM with sipper start are dropped.

```{r time_lookup}
minute_vals <- formatC(0:59, width = 2, flag = "0")

hour_vals <- c("10","11","12","13","14","15","16","17","18",
               "19","20","21","22","23","0","1","2","3",
               "4","5","6","7","8","9")

hour_min <- NA
for (x in seq_along(hour_vals))
  for (y in seq_along(minute_vals))
    hour_min <- c(hour_min, paste(hour_vals[x], minute_vals[y], sep = ":"))

hour_min         <- hour_min[!is.na(hour_min)]
rld_seq_time     <- seq(0, length(hour_min) - 1)
time_rldseq_df   <- data.frame(Time = hour_min, Time_Number = rld_seq_time)
```

---

## Paths

```{r paths}
main_dir   <- "C:/Users/rache/OneDrive - University of Pittsburgh/Documents/Lab/Rotation 2/Home Cage Sipper Experiment"
sipper_dir <- file.path(main_dir, "Sipper Data")

out_dir <- sipper_dir
for (d in c("binned_files", "results", "figures")) {
  dir.create(file.path(out_dir, d), showWarnings = FALSE)
}
```

---

# Data ingestion & 1-min binning

Note: Group metadata are derived from folder structure.

Expected tree:
    Sipper Data > [Sex] [Fluid] > [Age]_week > Day [N] [EtOH Side] [Mouse ID] [Date].CSV
      e.g., Sipper Data/Female EtOH/3_week/Day 2 EtOH Right 22 110721.CSV

  Sex = first token of the sipper_dir subfolder (Female / Male)
  Fluid group = second token of the sipper_dir subfolder (EtOH / H2O)
  Age = numeric prefix of the age folder, second subfolder (3 / 6 / 18)
  Experimental Day = V2 position in filename split
  Side of EtOH bottle = V4 position in filename split
  Sample = V5 position in filename split

```{r binning, results='hide'}
collect_files <- function(sipper_dir) {
  fluid_dirs <- list.dirs(sipper_dir, recursive = FALSE)
  out <- list()
  for (fd in fluid_dirs) {
    tokens    <- strsplit(basename(fd), " ")[[1]]
    sex_val   <- tokens[1]
    fluid_val <- tokens[2]
    age_dirs  <- list.dirs(fd, recursive = FALSE)
    for (ad in age_dirs) {
      age_val <- as.integer(sub("_week.*", "", basename(ad)))
      csvs    <- list.files(ad, pattern = "\\.CSV$", full.names = TRUE,
                            ignore.case = TRUE)
      if (length(csvs) == 0) next
      tmp             <- data.frame(FilePath = csvs, stringsAsFactors = FALSE)
      tmp$Sex         <- sex_val
      tmp$FluidGroup  <- fluid_val
      tmp$Age         <- age_val
      out[[length(out) + 1]] <- tmp
    }
  }
  do.call("rbind", out)
}

file_meta <- collect_files(sipper_dir)

fn_tokens <- strsplit(basename(file_meta$FilePath), " ")
file_meta$Sample    <- sapply(fn_tokens, function(x) if (length(x) >= 5) x[5] else NA)
file_meta$EtOH_Side <- sapply(fn_tokens, function(x) if (length(x) >= 4) x[4] else NA)
file_meta$ExpDay    <- sapply(fn_tokens, function(x) if (length(x) >= 2) as.numeric(x[2]) else NA)

n_before  <- nrow(file_meta)
file_meta <- file_meta[!is.na(file_meta$FilePath) &
                        !is.na(file_meta$Sample)   &
                        !is.na(file_meta$EtOH_Side) &
                        !is.na(file_meta$ExpDay), ]
n_dropped <- n_before - nrow(file_meta)
if (n_dropped > 0)
  cat("WARNING: dropped", n_dropped, "file(s) with unparseable names:\n")

file_meta$UniqueID <- paste(file_meta$Sex, file_meta$FluidGroup,
                             file_meta$Age, file_meta$Sample, sep = "_")

cat("\nParsed file metadata:\n")
print(file_meta[, c("FilePath","Sex","FluidGroup","Age","Sample","UniqueID","EtOH_Side","ExpDay")])

sample_ids <- unique(file_meta$UniqueID)

for (f in seq_along(sample_ids)) {

  samp <- sample_ids[f]

  if (is.na(samp)) {
    cat("Skipping unparseable sample ID at index", f, "\n")
    next
  }

  out_file  <- file.path(out_dir, "binned_files",
                         paste0("Sample_", samp, "_1min_bins_spf.txt"))

  if (file.exists(out_file)) {
    cat("Skipping already-processed sample:", samp, "\n")
    next
  }

  samp_meta    <- file_meta[file_meta$UniqueID == samp, ]
  sample_files <- samp_meta$FilePath
  sample_list  <- list()

  for (i in seq_along(sample_files)) {
    cat(i, ":", sample_files[i], "\n")
    file_in     <- read.csv(sample_files[i])
    new_file_df <- as.data.frame(matrix(nrow = nrow(file_in),
                                        ncol = ncol(file_in)))
    names(new_file_df) <- names(file_in)
    new_file_df[, 1]   <- file_in[, 1]
    new_file_df[, 2]   <- file_in[, 2]
    new_file_df[, 3]   <- file_in[, 3]
    new_file_df[, 8]   <- file_in[, 8]

    for (x in 2:nrow(file_in)) {
      for (col in c("LeftCount", "RightCount", "LeftDuration", "RightDuration")) {
        curr <- file_in[[col]][x]; prev <- file_in[[col]][x - 1]
        new_file_df[[col]][x] <- if (!is.na(curr) && !is.na(prev) && curr > prev)
          curr - prev else NA
      }
    }

    new_file_df[is.na(new_file_df)] <- 0

    split_s_d    <- strsplit(new_file_df$MM.DD.YYYY.hh.mm.ss, " ")
    split_s_d_df <- as.data.frame(do.call("rbind", split_s_d))
    names(split_s_d_df) <- c("Day", "Time")

    for (y in seq_len(nrow(split_s_d_df))) {
      spl_t <- strsplit(split_s_d_df$Time[y], ":")[[1]]
      split_s_d_df$Time[y] <- paste(spl_t[1:2], collapse = ":")
      spl_d <- strsplit(split_s_d_df$Day[y], "/")[[1]]
      split_s_d_df$Day[y]  <- paste(spl_d[1], spl_d[2], spl_d[3], sep = "/")
    }

    new_file_df             <- cbind(new_file_df, split_s_d_df)
    new_file_df$File_Name   <- sample_files[i]
    new_file_df$ExpDay      <- samp_meta$ExpDay[i]
    new_file_df$EtOH_Side   <- samp_meta$EtOH_Side[i]
    new_file_df$Sex         <- samp_meta$Sex[i]
    new_file_df$FluidGroup  <- samp_meta$FluidGroup[i]
    new_file_df$Age         <- samp_meta$Age[i]
    sample_list[[i]]        <- new_file_df
    rm(new_file_df)
  }

  sample_df           <- do.call("rbind", sample_list)
  sample_df$Sample    <- samp
  sample_df$MouseNum  <- file_meta$Sample[file_meta$UniqueID == samp][1]
  sample_df$ExpDay    <- as.numeric(sample_df$ExpDay)
  sample_df           <- sample_df[order(sample_df$ExpDay), ]

  day_table <- as.data.frame(table(sample_df$Day))
  day_list  <- list()

  for (z in seq_len(nrow(day_table))) {
    day_val    <- as.character(day_table$Var1[z])
    day_val_df <- sample_df[sample_df$Day == day_val, ]

    time_table_df             <- as.data.frame(table(day_val_df$Time))
    time_table_df$Device      <- as.character(unique(day_val_df$Device)[1])
    time_table_df$LeftCount   <- NA; time_table_df$LeftDuration   <- NA
    time_table_df$RightCount  <- NA; time_table_df$RightDuration  <- NA
    time_table_df$Day         <- day_val

    for (a in seq_len(nrow(time_table_df))) {
      tv        <- as.character(time_table_df$Var1[a])
      sub_day   <- day_val_df[day_val_df$Time == tv, ]
      time_table_df$LeftCount[a]     <- sum(sub_day$LeftCount)
      time_table_df$LeftDuration[a]  <- sum(sub_day$LeftDuration)
      time_table_df$RightCount[a]    <- sum(sub_day$RightCount)
      time_table_df$RightDuration[a] <- sum(sub_day$RightDuration)
    }

    names(time_table_df)[1]  <- "Time"
    nondup                   <- day_val_df[!duplicated(day_val_df$Time), ]
    nondup_short             <- nondup[, c("BatteryVoltage", "Time", "File_Name",
                                           "ExpDay", "EtOH_Side", "Sample",
                                           "Sex", "FluidGroup", "Age")]
    merge1 <- merge(time_table_df, nondup_short, by = "Time")
    merge2 <- merge(merge1, time_rldseq_df, by = "Time", all = TRUE)
    merge2 <- merge2[order(merge2$Time_Number), ]

    for (a in 2:nrow(merge2)) {
      if (is.na(merge2$Device[a])) {
        for (col in c("Device","Day","File_Name","ExpDay","EtOH_Side",
                      "Sample","Sex","FluidGroup","Age"))
          merge2[[col]][a] <- merge2[[col]][a - 1]
      }
    }
    if (any(is.na(merge2$Device))) {
      first_ok <- which(!is.na(merge2$Device))[1]
      for (col in c("Device","Day","File_Name","ExpDay","EtOH_Side",
                    "Sample","Sex","FluidGroup","Age"))
        merge2[[col]][is.na(merge2[[col]])] <- merge2[[col]][first_ok]
    }

    day_list[[z]] <- merge2
  }

  day_df <- do.call("rbind", day_list)
  day_df <- day_df[!is.na(day_df$Time_Number), ]

  day_df$H2O_Count    <- NA; day_df$H2O_Duration <- NA
  day_df$EtOH_Count   <- NA; day_df$EtOH_Duration <- NA

  for (x in seq_len(nrow(day_df))) {
    if (day_df$EtOH_Side[x] == "Right") {
      day_df$H2O_Count[x]    <- day_df$LeftCount[x]
      day_df$H2O_Duration[x] <- day_df$LeftDuration[x]
      day_df$EtOH_Count[x]   <- day_df$RightCount[x]
      day_df$EtOH_Duration[x]<- day_df$RightDuration[x]
    } else {
      day_df$H2O_Count[x]    <- day_df$RightCount[x]
      day_df$H2O_Duration[x] <- day_df$RightDuration[x]
      day_df$EtOH_Count[x]   <- day_df$LeftCount[x]
      day_df$EtOH_Duration[x]<- day_df$LeftDuration[x]
    }
  }

  if (unique(day_df$FluidGroup)[1] == "H2O") {
    day_df$H2O_Count    <- day_df$LeftCount    + day_df$RightCount
    day_df$H2O_Duration <- day_df$LeftDuration + day_df$RightDuration
    day_df$EtOH_Count   <- 0
    day_df$EtOH_Duration<- 0
  }

  day_df$Total_Count    <- day_df$H2O_Count    + day_df$EtOH_Count
  day_df$Total_Duration <- day_df$H2O_Duration + day_df$EtOH_Duration

  day_df$Cycle <- ifelse(day_df$Time_Number < 720, "Dark", "Light")

  write.table(day_df,
              file      = out_file,
              row.names = FALSE, quote = FALSE, sep = "\t")
  cat("Written:", out_file, "\n")
}
```

---

# 60-min binning

```{r sixty_min_bins, results='hide'}
setwd(file.path(out_dir, "binned_files"))
one_min_files   <- list.files(pattern = "_1min_bins_spf\\.txt$", full.names = TRUE)
allsample_df    <- as.data.frame(do.call("rbind", lapply(one_min_files, read.delim,
                                                    stringsAsFactors = FALSE)))
allsample_df$ExpDay <- as.numeric(allsample_df$ExpDay)

processed_list <- list()

for (x in seq_along(one_min_files)) {
  indiv_file  <- read.delim(one_min_files[x], stringsAsFactors = FALSE)
  unique_date <- unique(indiv_file$Day)
  date_list   <- list()

  for (y in seq_along(unique_date)) {
    sub_date_df           <- indiv_file[indiv_file$Day == unique_date[y], ]
    sub_date_df$Time_Vals <- seq_len(nrow(sub_date_df))
    bin_val               <- 60
    bin_seq               <- seq(1, 1440, by = bin_val)
    if (max(bin_seq) != 1440) bin_seq <- c(bin_seq, 1440)
    temp_list <- list()
    a <- 1
    while (a < length(bin_seq)) {
      end_idx  <- if (a + 1 != length(bin_seq)) bin_seq[a + 1] - 1 else bin_seq[a + 1]
      time_sub <- sub_date_df[bin_seq[a]:end_idx, ]
      sum_cols <- intersect(c("LeftCount","LeftDuration","RightCount","RightDuration",
                              "H2O_Count","H2O_Duration","EtOH_Count","EtOH_Duration",
                              "Total_Count","Total_Duration"),
                            names(time_sub))
      summed           <- colSums(time_sub[, sum_cols], na.rm = TRUE)
      trim_row         <- time_sub[1, ]
      trim_row[, sum_cols] <- summed
      trim_row$Bin_Time    <- "60 min"
      temp_list[[a]]       <- trim_row
      a <- a + 1
    }
    date_list[[y]] <- do.call("rbind", temp_list)
  }
  processed_list[[x]] <- do.call("rbind", date_list)
}

for (x in seq_along(processed_list)) {
  proc_df  <- processed_list[[x]]
  samp_num <- proc_df$Sample[1]
  out_file <- file.path(out_dir, "binned_files",
                        paste0("Sample_", samp_num, "_binned_values_spf.txt"))
  if (!file.exists(out_file)) {
    write.table(proc_df, file = out_file, row.names = FALSE, quote = FALSE, sep = "\t")
    cat("Written:", out_file, "\n")
  }
}

binned_dataset <- as.data.frame(do.call("rbind", processed_list))
binned_dataset$Age        <- as.factor(as.character(binned_dataset$Age))
binned_dataset$Sex        <- as.factor(as.character(binned_dataset$Sex))
binned_dataset$FluidGroup <- as.factor(as.character(binned_dataset$FluidGroup))
setwd(main_dir)
```

---

# Shared circadian plot utilities

```{r plot_utils}
age_colors <- c("3"  = "#3E067D",
                "6"  = "#F70065",
                "18" = "#277f80")

age_labels <- c("3" = "3 wk", "6" = "6 wk", "18" = "18 wk")

shared_x_noon <- scale_x_continuous(
  breaks = seq(0, 1440, by = 120),
  labels = c("10:00","12:00","14:00","16:00","18:00","20:00","22:00",
             "00:00","02:00","04:00","06:00","08:00","09:59"),
  name   = NULL)

shared_theme_circ <- theme_minimal() +
    theme(
      axis.text.x  = element_text(angle = 45, hjust = 1, face = "bold", size = 12),
      axis.text.y  = element_text(face = "bold", size = 12),
      axis.title.x = element_text(face = "bold", size = 14),
      axis.title.y = element_text(face = "bold", size = 14, vjust = 3),
      plot.title   = element_text(face = "bold", size = 16),
      legend.title = element_text(face = "bold", size = 14),
      legend.text  = element_text(face = "bold", size = 12),
      strip.text   = element_text(face = "bold", size = 14),
      text         = element_text(size = 10),
      panel.spacing = unit(0.3, "lines"))

dark_rect_noon <- annotate("rect", xmin = -Inf, xmax = 720,
                           ymin = -Inf, ymax = Inf,
                           fill = "#939193", alpha = 0.2)
```

---

# Circadian plots

## EtOH group

```{r circ_etoh_group, fig.width=14, fig.height=6, results='asis'}

make_circ_plot <- function(sum_df, measure_col, ylab_text,
                           title_str, y_max = 200) {
  sum_df$Age <- factor(sum_df$Age, levels = c("3","6","18"))
  se_df      <- sum_df[!is.na(sum_df$se), ]

  p <- ggplot(sum_df, aes(x = Time_Number, y = Mean,
                          fill = Age, color = Age)) +
    dark_rect_noon +
    geom_bar(stat = "identity", alpha = 0.7) +
    geom_errorbar(data = se_df,
                  aes(ymin = Mean - se, ymax = Mean + se),
                  width = 0.5, color = "black", linewidth = 0.3) +
    geom_smooth(se = FALSE, linewidth = 0.8, color = "black") +
    scale_fill_manual(values = age_colors, labels = age_labels, name = "Age") +
    scale_color_manual(values = age_colors, labels = age_labels, name = "Age") +
    facet_wrap(~ Age, nrow = 1, labeller = labeller(Age = age_labels)) +
    shared_x_noon + shared_theme_circ +
    ylab(ylab_text) + ggtitle(title_str) +
    coord_cartesian(ylim = c(0, y_max))
  p
}

etoh_binned <- binned_dataset[binned_dataset$FluidGroup == "EtOH" &
                                binned_dataset$Bin_Time  == "60 min", ]

for (sex_val in c("Female", "Male")) {

  sub_sex <- etoh_binned[etoh_binned$Sex == sex_val, ]

  for (measure in c("EtOH_Count", "H2O_Count", "Total_Count")) {

    tab_label <- switch(measure,
                        EtOH_Count  = "EtOH consumption",
                        H2O_Count   = "H2O consumption",
                        Total_Count = "Total consumption")

    pa <- aggregate(as.formula(paste(measure, "~ Sample + Time_Number + Age")),
                    data = sub_sex, FUN = mean, na.rm = TRUE)
    sm <- summarySE(pa, measurevar = measure,
                    groupvars = c("Time_Number", "Age"), na.rm = TRUE)
    names(sm)[names(sm) == "mean"] <- "Mean"

    y_max <- if (measure == "Total_Count") 300 else 200

    p <- make_circ_plot(
      sum_df      = sm,
      measure_col = measure,
      ylab_text   = paste("Mean", gsub("_", " ", measure), "/ 60 min"),
      title_str   = paste(tab_label, "—", sex_val, "— EtOH group"),
      y_max       = y_max
    )
    print(p)

    ggsave(
      filename = file.path(out_dir, "figures",
                           paste0("Circ_EtOH_", sex_val, "_", measure, ".pdf")),
      plot = p, width = 14, height = 5, units = "in"
    )
  }
}
```

## H2O group

```{r circ_h2o_group, fig.width=14, fig.height=6, results='asis'}

h2o_binned <- binned_dataset[binned_dataset$FluidGroup == "H2O" &
                               binned_dataset$Bin_Time  == "60 min", ]

for (sex_val in c("Female", "Male")) {
  sub_sex <- h2o_binned[h2o_binned$Sex == sex_val, ]

  pa <- aggregate(H2O_Count ~ Sample + Time_Number + Age,
                  data = sub_sex, FUN = mean, na.rm = TRUE)
  sm <- summarySE(pa, measurevar = "H2O_Count",
                  groupvars = c("Time_Number", "Age"), na.rm = TRUE)
  names(sm)[names(sm) == "mean"] <- "Mean"
  p <- make_circ_plot(
    sum_df      = sm,
    measure_col = "H2O_Count",
    ylab_text   = "Mean H2O Counts / 60 min",
    title_str   = paste("H2O consumption —", sex_val, "— H2O group"),
    y_max       = 350
  )
  print(p)

  ggsave(
    filename = file.path(out_dir, "figures",
                         paste0("Circ_H2O_", sex_val, "_H2O_Count.pdf")),
    plot = p, width = 14, height = 5, units = "in"
  )
}
```

---

# Per-animal QC plots

```{r qc_plots, results='hide'}
setwd(file.path(out_dir, "binned_files"))
one_min_bin_files <- list.files(pattern = "_1min_bins_spf\\.txt$")

pdf(file = file.path(out_dir, "figures", "QC_one_min_bin_plots.pdf"),
    height = 8.5, width = 14)

for (x in seq_along(one_min_bin_files)) {
  day_df <- read.delim(one_min_bin_files[x], stringsAsFactors = FALSE)
  day_df$ExpDay <- factor(day_df$ExpDay,
                          levels = unique(sort(day_df$ExpDay)))

  max_y <- max(c(day_df$EtOH_Count, day_df$H2O_Count), na.rm = TRUE)

  p_etoh <- ggplot(day_df, aes(x = Time_Number, y = EtOH_Count, fill = ExpDay)) +
    annotate("rect", xmin = -Inf, xmax = 720, ymin = -Inf, ymax = Inf,
             fill = "darkgrey", alpha = 0.1) +
    geom_bar(stat = "identity") + theme_bw() + shared_x_noon +
    ggtitle(paste("Sample", unique(day_df$Sample), "— EtOH")) +
    ylab("EtOH Counts") + ylim(0, max_y) +
    facet_grid(ExpDay ~ ., scales = "fixed") +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

  p_h2o <- ggplot(day_df, aes(x = Time_Number, y = H2O_Count, fill = ExpDay)) +
    annotate("rect", xmin = -Inf, xmax = 720, ymin = -Inf, ymax = Inf,
             fill = "darkgrey", alpha = 0.1) +
    geom_bar(stat = "identity") + theme_bw() + shared_x_noon +
    ggtitle(paste("Sample", unique(day_df$Sample), "— H2O")) +
    ylab("H2O Counts") + ylim(0, max_y) +
    facet_grid(ExpDay ~ ., scales = "fixed") +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

  print(p_etoh + p_h2o)
}
dev.off()
setwd(main_dir)
```

---

# Cosinor modelling

Interested in determining and comparing across age/sex/fluid group:
  - Acrophase (approx. peak of rhythm)
  - Amplitude (magnitude of rhythm)
  - MESOR (mean value)

## Fit model

```{r cosinor_fit}
tn_to_clock <- function(tn) {
  total_min <- as.integer(tn %% 1440)
  total_min_from_midnight <- as.integer((10L * 60L + total_min) %% 1440L)
  h    <- total_min_from_midnight %/% 60L
  m    <- total_min_from_midnight %%  60L
  ampm <- ifelse(h < 12L, "AM", "PM")
  h12  <- ifelse(h %% 12L == 0L, 12L, h %% 12L)
  paste0(h12, ":", formatC(m, width = 2, flag = "0"), " ", ampm)
}

fit_cosinor <- function(time_vec, count_vec) {
  df <- data.frame(x = time_vec, y = count_vec)
  df <- df[!is.na(df$y) & !is.na(df$x), ]
  if (nrow(df) < 6) return(NULL)
  w       <- 2 * pi / 1440
  df$cosx <- cos(w * df$x)
  df$sinx <- sin(w * df$x)
  fit <- tryCatch(lm(y ~ cosx + sinx, data = df), error = function(e) NULL)
  if (is.null(fit)) return(NULL)
  cf    <- coef(fit)
  beta  <- cf["cosx"]; gamma <- cf["sinx"]
  amp   <- sqrt(beta^2 + gamma^2)

  t_seq  <- seq(0, 1439, by = 1)
  fitted <- unname(cf["(Intercept)"]) + beta * cos(w * t_seq) + gamma * sin(w * t_seq)
  phi_tn <- t_seq[which.max(fitted)]

  data.frame(
    MESOR           = unname(cf["(Intercept)"]),
    Amplitude       = unname(amp),
    Acrophase       = phi_tn,
    Acrophase_Clock = tn_to_clock(phi_tn),
    R2              = summary(fit)$r.squared
  )
}

bin60 <- binned_dataset[binned_dataset$Bin_Time == "60 min", ]

cosinor_measures <- list(
  EtOH = c("EtOH_Count", "H2O_Count", "Total_Count"),
  H2O  = c("Total_Count")
)

cosinor_list <- list()

for (fg in c("EtOH", "H2O")) {
  bin_fg <- bin60[bin60$FluidGroup == fg, ]
  for (var in cosinor_measures[[fg]]) {
    pa <- aggregate(as.formula(paste(var, "~ Sample + Sex + Age + Time_Number")),
                    data = bin_fg, FUN = mean, na.rm = TRUE)
    for (s in unique(pa$Sample)) {
      sub <- pa[pa$Sample == s, ]
      res <- fit_cosinor(sub$Time_Number, sub[[var]])
      if (!is.null(res)) {
        res$Sample     <- s
        res$Sex        <- sub$Sex[1]
        res$Age        <- sub$Age[1]
        res$FluidGroup <- fg
        res$Measure    <- var
        cosinor_list[[paste(fg, var, s)]] <- res
      }
    }
  }
}

cosinor_df     <- as.data.frame(do.call("rbind", cosinor_list))
cosinor_df$Age <- factor(cosinor_df$Age, levels = c("3","6","18"))
cosinor_df$Sex <- factor(cosinor_df$Sex, levels = c("Female","Male"))

write.csv(cosinor_df,
          file = file.path(out_dir, "results", "Cosinor_PerAnimal.csv"),
          row.names = FALSE, quote = FALSE)

print(cosinor_df[, c("Sample","Sex","Age","FluidGroup","Measure",
                      "MESOR","Amplitude","Acrophase","Acrophase_Clock","R2")])
```


## Plot

```{r cosinor_plots, fig.width=14, fig.height=5, results='asis'}

make_cosinor_panel <- function(df, param, y_label, title_str) {
  df$Age <- factor(df$Age, levels = c("3","6","18"))
  y_max  <- switch(param, Acrophase = 500, Amplitude = 150, MESOR = 200)
  sm <- aggregate(as.formula(paste(param, "~ Age + Sex")),
                  data = df, FUN = mean, na.rm = TRUE)
  se <- aggregate(as.formula(paste(param, "~ Age + Sex")),
                  data = df,
                  FUN = function(x) sd(x, na.rm=TRUE) / sqrt(sum(!is.na(x))))
  names(sm)[names(sm) == param] <- "Mean"
  names(se)[names(se) == param] <- "SE"
  sm <- merge(sm, se, by = c("Age","Sex"))
  ggplot(sm, aes(x = Age, y = Mean, fill = Age)) +
    geom_bar(stat = "identity", width = 0.6, linewidth = 0.7,
             color = "black", alpha = 0.7) +
    geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE),
                  width = 0.15, linewidth = 0.7, color = "black") +
    geom_jitter(data = df, aes(x = Age, y = .data[[param]], fill = Age),
                shape = 21, size = 2, width = 0.1,
                stroke = 0.7, color = "black", alpha = 0.7) +
    scale_fill_manual(values = age_colors, labels = age_labels, name = "Age") +
    scale_x_discrete(labels = age_labels) +
    coord_cartesian(ylim = c(0, y_max)) +
    facet_wrap(~ Sex) +
    ylab(y_label) + xlab("Age") + ggtitle(title_str) +
    theme_minimal() +
    theme(
      axis.text.x  = element_text(angle = 45, hjust = 1, face = "bold", size = 12),
      axis.text.y  = element_text(face = "bold", size = 12),
      axis.title.x = element_text(face = "bold", size = 14),
      axis.title.y = element_text(face = "bold", size = 14, vjust = 3),
      plot.title   = element_text(face = "bold", size = 16),
      legend.position = "none",
      strip.text   = element_text(face = "bold", size = 14),
      text         = element_text(size = 10),
      panel.spacing = unit(0.3, "lines"))
}

for (fg in c("EtOH","H2O")) {
  cos_fg <- cosinor_df[cosinor_df$FluidGroup == fg, ]
  for (var in cosinor_measures[[fg]]) {
    cos_m  <- cos_fg[cos_fg$Measure == var, ]
    p_acro <- make_cosinor_panel(cos_m, "Acrophase", "Acrophase (min from 10AM)",
                                  paste("Peak Time —", var, "—", fg))
    p_amp  <- make_cosinor_panel(cos_m, "Amplitude",  "Amplitude (counts/60min)",
                                  paste("Amplitude —", var, "—", fg))
    p_mes  <- make_cosinor_panel(cos_m, "MESOR",      "MESOR (counts/60min)",
                                  paste("MESOR —", var, "—", fg))
    print(p_acro | p_amp | p_mes)
    ggsave(filename = file.path(out_dir, "figures",
                                paste0("Cosinor_", fg, "_", var, ".pdf")),
           plot = p_acro | p_amp | p_mes, width = 14, height = 5, units = "in")
  }
}
```

## Stats

```{r cosinor_stats, results='asis'}
cosinor_stat_rows <- list()
for (fg in c("EtOH", "H2O")) {
  cos_fg <- cosinor_df[cosinor_df$FluidGroup == fg, ]
  for (var in cosinor_measures[[fg]]) {
    cos_m <- cos_fg[cos_fg$Measure == var, ]
    for (param in c("Acrophase", "Amplitude", "MESOR")) {
      cos_m$DV  <- cos_m[[param]]
      cos_m$Age <- factor(cos_m$Age, levels = c("3","6","18"))
      cos_m$Sex <- factor(cos_m$Sex, levels = c("Female","Male"))
      cell_n <- cos_m %>%
        group_by(Age, Sex) %>%
        summarise(n = n_distinct(Sample), .groups = "drop")

      res <- tryCatch(
        aov(DV ~ Age * Sex, data = cos_m),
        error = function(e) { cat("ERROR:", conditionMessage(e), "\n"); NULL }
      )
      if (!is.null(res)) {
        s   <- summary(res)
        tbl <- as.data.frame(s[[1]])
        tbl$Effect     <- rownames(tbl)
        tbl$FluidGroup <- fg
        tbl$Measure    <- var
        tbl$Parameter  <- param
        names(tbl)     <- make.names(names(tbl))
        if ("Pr..F." %in% names(tbl)) tbl$p.value <- tbl[["Pr..F."]]
        tbl$sig <- ifelse(tbl$p.value < 0.001, "***",
                   ifelse(tbl$p.value < 0.01,  "**",
                   ifelse(tbl$p.value < 0.05,  "*",
                   ifelse(tbl$p.value < 0.1,   ".", ""))))
        cosinor_stat_rows[[paste(fg, var, param)]] <- tbl
      }
    }  
  }   
}     

if (length(cosinor_stat_rows) > 0) {
  tidy_cos <- do.call("rbind", lapply(cosinor_stat_rows, function(tbl) {
    tbl  <- as.data.frame(tbl)
    core <- c("FluidGroup","Measure","Parameter","Effect",
              "Df","Sum.Sq","Mean.Sq","F.value","p.value","sig")
    tbl[, intersect(core, names(tbl)), drop = FALSE]
  }))
  tidy_cos[sapply(tidy_cos, is.numeric)] <-
    lapply(tidy_cos[sapply(tidy_cos, is.numeric)],
           function(x) round(as.numeric(x), 4))
  write.csv(tidy_cos,
            file      = file.path(out_dir, "results", "Cosinor_Stats_Summary.csv"),
            row.names = FALSE, quote = FALSE)
}
```

There were significant Age x Sex interactions for acrophase for EtOH, EtOH Total, & H2O.
Perform Tukey post-hoc analyses:

```{r cosinor_tukey_etoh, results='asis'}

tukey_rows_etoh <- list()

for (var in c("EtOH_Count", "H2O_Count", "Total_Count")) {
  cos_m <- cosinor_df[cosinor_df$FluidGroup == "EtOH" &
                        cosinor_df$Measure    == var, ]
  cos_m$Age <- factor(cos_m$Age, levels = c("3","6","18"))
  cos_m$Sex <- factor(cos_m$Sex, levels = c("Female","Male"))

  res <- tryCatch(
    aov(Acrophase ~ Age * Sex, data = cos_m),
    error = function(e) { cat("ERROR:", conditionMessage(e), "\n"); NULL }
  )

  if (!is.null(res)) {
    tukey <- TukeyHSD(res, which = "Age:Sex")
    tbl             <- as.data.frame(tukey$`Age:Sex`)
    tbl$Comparison  <- rownames(tbl)
    tbl$Measure     <- var
    tbl$FluidGroup  <- "EtOH"
    names(tbl)[1:4] <- c("Difference","Lower_CI","Upper_CI","p.adj")
    tbl$sig <- ifelse(tbl$p.adj < 0.001, "***",
               ifelse(tbl$p.adj < 0.01,  "**",
               ifelse(tbl$p.adj < 0.05,  "*",
               ifelse(tbl$p.adj < 0.1,   ".", ""))))
    tbl <- tbl[order(tbl$p.adj), ]
    tbl[, c("Difference","Lower_CI","Upper_CI","p.adj")] <-
      lapply(tbl[, c("Difference","Lower_CI","Upper_CI","p.adj")],
             function(x) round(x, 4))
    tukey_rows_etoh[[var]] <- tbl
  }
}

if (length(tukey_rows_etoh) > 0) {
  tukey_etoh_df <- do.call("rbind", tukey_rows_etoh)
  write.csv(tukey_etoh_df,
            file      = file.path(out_dir, "results",
                                  "Tukey_Acrophase_AgeSex_EtOHgroup.csv"),
            row.names = FALSE, quote = FALSE)
}
```

```{r cosinor_tukey_h2o, results='asis'}

cos_m <- cosinor_df[cosinor_df$FluidGroup == "H2O" &
                      cosinor_df$Measure    == "Total_Count", ]
cos_m$Age <- factor(cos_m$Age, levels = c("3","6","18"))
cos_m$Sex <- factor(cos_m$Sex, levels = c("Female","Male"))

res <- tryCatch(
  aov(Acrophase ~ Age * Sex, data = cos_m),
  error = function(e) { cat("ERROR:", conditionMessage(e), "\n"); NULL }
)

if (!is.null(res)) {
  tukey <- TukeyHSD(res, which = "Age:Sex")
  tbl             <- as.data.frame(tukey$`Age:Sex`)
  tbl$Comparison  <- rownames(tbl)
  tbl$Measure     <- "Total_Count"
  tbl$FluidGroup  <- "H2O"
  names(tbl)[1:4] <- c("Difference","Lower_CI","Upper_CI","p.adj")
  tbl$sig <- ifelse(tbl$p.adj < 0.001, "***",
             ifelse(tbl$p.adj < 0.01,  "**",
             ifelse(tbl$p.adj < 0.05,  "*",
             ifelse(tbl$p.adj < 0.1,   ".", ""))))
  tbl <- tbl[order(tbl$p.adj), ]
  tbl[, c("Difference","Lower_CI","Upper_CI","p.adj")] <-
    lapply(tbl[, c("Difference","Lower_CI","Upper_CI","p.adj")],
           function(x) round(x, 4))

  write.csv(tbl,
            file      = file.path(out_dir, "results",
                                  "Tukey_Acrophase_AgeSex_H2Ogroup.csv"),
            row.names = FALSE, quote = FALSE)
}
```

---

# Directly compare Cosinor for total fluid across fluid groups

```{r cosinor_fluid_compare_plot}

cos_total <- cosinor_df[cosinor_df$Measure == "Total_Count", ]
cos_total$Age        <- factor(cos_total$Age,        levels = c("3","6","18"))
cos_total$Sex        <- factor(cos_total$Sex,        levels = c("Female","Male"))
cos_total$FluidGroup <- factor(cos_total$FluidGroup, levels = c("H2O","EtOH"))

fluid_alpha <- c("EtOH" = 0.9, "H2O" = 0.4)

make_fluid_comparison_panel <- function(df, param, y_label, title_str) {

  y_max <- switch(param,
                  Acrophase = 500,
                  Amplitude = 150,
                  MESOR     = 200)

  sm <- aggregate(as.formula(paste(param, "~ Age + Sex + FluidGroup")),
                  data = df, FUN = mean, na.rm = TRUE)
  se <- aggregate(as.formula(paste(param, "~ Age + Sex + FluidGroup")),
                  data = df,
                  FUN = function(x) sd(x, na.rm=TRUE) / sqrt(sum(!is.na(x))))
  names(sm)[names(sm) == param] <- "Mean"
  names(se)[names(se) == param] <- "SE"
  sm <- merge(sm, se, by = c("Age","Sex","FluidGroup"))

  ggplot(sm, aes(x = Age, y = Mean, fill = Age, group = FluidGroup)) +
    geom_bar(aes(alpha = FluidGroup), stat = "identity", width = 0.6,
             linewidth = 0.7, color = "black",
             position = position_dodge(width = 0.7)) +
    geom_jitter(data = df,
                aes(x = Age, y = .data[[param]], fill = Age,
                    alpha = FluidGroup),
                shape = 21, size = 2, stroke = 0.7, color = "black",
                position = position_jitterdodge(jitter.width = 0.1,
                                                dodge.width  = 0.7)) +
    geom_errorbar(aes(ymin = Mean - SE, ymax = Mean + SE),
                  width = 0.15, linewidth = 0.7, color = "black",
                  position = position_dodge(width = 0.7)) +
    scale_fill_manual(values = age_colors, labels = age_labels, name = "Age") +
    scale_alpha_manual(values = fluid_alpha, name = "Fluid group") +
    scale_x_discrete(labels = age_labels) +
    coord_cartesian(ylim = c(0, y_max)) +
    facet_wrap(~ Sex) +
    ylab(y_label) + xlab("Age") + ggtitle(title_str) +
    theme_minimal() +
    theme(
      axis.text.x  = element_text(angle = 45, hjust = 1, face = "bold", size = 12),
      axis.text.y  = element_text(face = "bold", size = 12),
      axis.title.x = element_text(face = "bold", size = 14),
      axis.title.y = element_text(face = "bold", size = 14, vjust = 3),
      plot.title   = element_text(face = "bold", size = 16),
      legend.position = "none",
      strip.text   = element_text(face = "bold", size = 14),
      text         = element_text(size = 10),
      panel.spacing = unit(0.3, "lines"))
}

p_acro <- make_fluid_comparison_panel(cos_total, "Acrophase",
                                       "Acrophase (min from 10AM)",
                                       "Peak Time — Total Consumption")
p_amp  <- make_fluid_comparison_panel(cos_total, "Amplitude",
                                       "Amplitude (counts/60min)",
                                       "Amplitude — Total Consumption")
p_mes  <- make_fluid_comparison_panel(cos_total, "MESOR",
                                       "MESOR (counts/60min)",
                                       "MESOR — Total Consumption")

print(p_acro | p_amp | p_mes)

ggsave(
  filename = file.path(out_dir, "figures", "Cosinor_FluidComparison_Total.pdf"),
  plot     = p_acro | p_amp | p_mes,
  width    = 16, height = 5, units = "in"
)
```

```{r cosinor_fluid_compare_stats}

fluid_stat_rows <- list()

for (sex_val in c("Female", "Male")) {
  cos_sex <- cos_total[cos_total$Sex == sex_val, ]

  for (param in c("Acrophase", "Amplitude", "MESOR")) {
    cos_sex$DV <- cos_sex[[param]]

    cell_n <- cos_sex %>%
      group_by(Age, FluidGroup) %>%
      summarise(n = n_distinct(Sample), .groups = "drop")

    if (any(cell_n$n < 2)) {
      cat("WARNING: cell(s) with < 2 animals — ANOVA skipped.\n\n")
      next
    }

    res <- tryCatch(
      aov(DV ~ Age * FluidGroup, data = cos_sex),
      error = function(e) { cat("ERROR:", conditionMessage(e), "\n"); NULL }
    )

    if (!is.null(res)) {
      s   <- summary(res)
      tbl <- as.data.frame(s[[1]])
      tbl$Effect    <- rownames(tbl)
      tbl$Parameter <- param
      tbl$Sex       <- sex_val
      names(tbl)    <- make.names(names(tbl))
      if ("Pr..F." %in% names(tbl)) tbl$p.value <- tbl[["Pr..F."]]
      tbl$sig <- ifelse(tbl$p.value < 0.001, "***",
                 ifelse(tbl$p.value < 0.01,  "**",
                 ifelse(tbl$p.value < 0.05,  "*",
                 ifelse(tbl$p.value < 0.1,   ".", ""))))
      fluid_stat_rows[[paste(sex_val, param)]] <- tbl
    }
  }
}

if (length(fluid_stat_rows) > 0) {
  tidy_fluid <- do.call("rbind", lapply(fluid_stat_rows, function(tbl) {
    tbl <- as.data.frame(tbl)
    core <- c("Sex","Parameter","Effect","Df","Sum.Sq","Mean.Sq","F.value","p.value","sig")
    tbl[, intersect(core, names(tbl)), drop = FALSE]
  }))
  tidy_fluid[sapply(tidy_fluid, is.numeric)] <-
    lapply(tidy_fluid[sapply(tidy_fluid, is.numeric)],
           function(x) round(as.numeric(x), 4))
  write.csv(tidy_fluid,
            file      = file.path(out_dir, "results",
                                  "Cosinor_FluidComparison_Stats.csv"),
            row.names = FALSE, quote = FALSE)
}
```

---

# Circadian overlay plots with acrophase markers

```{r circ_overlay, fig.width=8, fig.height=5, results='asis'}
make_circ_overlay <- function(sum_df, measure_col, ylab_text,
                              title_str, y_max = 200,
                              acro_df = NULL,
                              cosinor_params = NULL) {
  sum_df$Age <- factor(sum_df$Age, levels = c("3","6","18"))

  p <- ggplot(sum_df, aes(x = Time_Number, y = Mean, color = Age)) +
    annotate("rect", xmin = -Inf, xmax = 720,
             ymin = -Inf, ymax = Inf,
             fill = "#939193", alpha = 0.2) +
    scale_color_manual(values = age_colors, labels = age_labels, name = "Age") +
    shared_x_noon + shared_theme_circ +
    ylab(ylab_text) + ggtitle(title_str) +
    coord_cartesian(ylim = c(0, y_max)) +
    theme(legend.position = "right")

  if (!is.null(cosinor_params)) {
    cosinor_params$Age <- factor(cosinor_params$Age, levels = c("3","6","18"))
    w     <- 2 * pi / 1440
    t_seq <- seq(0, 1439, by = 1)
    curve_list <- lapply(1:nrow(cosinor_params), function(i) {
      row <- cosinor_params[i, ]
      data.frame(
        Time_Number = t_seq,
        Mean        = row$MESOR + row$Amplitude * cos(w * t_seq - w * row$Acrophase),
        Age         = row$Age
      )
    })
    curve_df     <- do.call("rbind", curve_list)
    curve_df$Age <- factor(curve_df$Age, levels = c("3","6","18"))
    p <- p + geom_line(data        = curve_df,
                       aes(x       = Time_Number,
                           y       = Mean,
                           color   = Age),
                       linewidth   = 1.2,
                       inherit.aes = FALSE)
  }

  if (!is.null(acro_df)) {
    acro_df$Age <- factor(acro_df$Age, levels = c("3","6","18"))
    p <- p +
      geom_jitter(data        = acro_df,
                 aes(x       = Acrophase,
                     y       = y_max * 0.95,
                     color   = Age),
                 shape       = 6,
                 size        = 3,
                 stroke      = 1,
                 position = position_jitter(height = 2, width = 2),
                 inherit.aes = FALSE)
  }
  p
}

etoh_binned <- binned_dataset[binned_dataset$FluidGroup == "EtOH" &
                                binned_dataset$Bin_Time  == "60 min", ]

for (sex_val in c("Female", "Male")) {
  sub_sex <- etoh_binned[etoh_binned$Sex == sex_val, ]

  for (measure in c("EtOH_Count", "H2O_Count", "Total_Count")) {
    tab_label <- switch(measure,
                        EtOH_Count  = "EtOH consumption",
                        H2O_Count   = "H2O consumption",
                        Total_Count = "Total consumption")

    pa <- aggregate(as.formula(paste(measure, "~ Sample + Time_Number + Age")),
                    data = sub_sex, FUN = mean, na.rm = TRUE)
    sm <- summarySE(pa, measurevar = measure,
                    groupvars = c("Time_Number", "Age"), na.rm = TRUE)
    names(sm)[names(sm) == "mean"] <- "Mean"
    y_max <- 250

    acro_sub   <- cosinor_df[cosinor_df$FluidGroup == "EtOH" &
                               cosinor_df$Measure    == measure &
                               cosinor_df$Sex        == sex_val, ]
    acro_means <- aggregate(cbind(Acrophase, Amplitude, MESOR) ~ Age,
                            data = acro_sub, FUN = mean, na.rm = TRUE)

    p_overlay <- make_circ_overlay(
      sum_df         = sm,
      measure_col    = measure,
      ylab_text      = paste("Mean", gsub("_", " ", measure), "/ 60 min"),
      title_str      = paste(tab_label, "—", sex_val, "— EtOH group"),
      y_max          = y_max,
      acro_df        = acro_means,
      cosinor_params = acro_means
    )
    print(p_overlay)
    ggsave(
      filename = file.path(out_dir, "figures",
                           paste0("Circ_EtOH_overlay_", sex_val, "_", measure, ".pdf")),
      plot = p_overlay, width = 8, height = 5, units = "in"
    )
  }
}

h2o_binned <- binned_dataset[binned_dataset$FluidGroup == "H2O" &
                               binned_dataset$Bin_Time  == "60 min", ]

for (sex_val in c("Female", "Male")) {
  sub_sex <- h2o_binned[h2o_binned$Sex == sex_val, ]

  pa <- aggregate(Total_Count ~ Sample + Time_Number + Age,
                  data = sub_sex, FUN = mean, na.rm = TRUE)
  sm <- summarySE(pa, measurevar = "Total_Count",
                  groupvars = c("Time_Number", "Age"), na.rm = TRUE)
  names(sm)[names(sm) == "mean"] <- "Mean"

  acro_sub   <- cosinor_df[cosinor_df$FluidGroup == "H2O" &
                             cosinor_df$Measure    == "Total_Count" &
                             cosinor_df$Sex        == sex_val, ]
  acro_means <- aggregate(cbind(Acrophase, Amplitude, MESOR) ~ Age,
                          data = acro_sub, FUN = mean, na.rm = TRUE)

  p_overlay <- make_circ_overlay(
    sum_df         = sm,
    measure_col    = "Total_Count",
    ylab_text      = "Mean Total Counts / 60 min",
    title_str      = paste("Total consumption —", sex_val, "— H2O group"),
    y_max          = 250,
    acro_df        = acro_means,
    cosinor_params = acro_means
  )
  print(p_overlay)
  ggsave(
    filename = file.path(out_dir, "figures",
                         paste0("Circ_H2O_overlay_", sex_val, "_Total_Count.pdf")),
    plot = p_overlay, width = 8, height = 5, units = "in"
  )
}
```

---